---
title: "Data Cleaning"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    #number_section: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(wooldridge)
library(palmerpenguins)
```

## Get Ready
### Download R and RStudio

Choose the one based on your own OS from https://posit.co/download/rstudio-desktop/.  
*Do not include space in the path.

[RCloud](https://posit.cloud/) can also be used.


### Tidyverse Package

R includes **R Basics** by its developers and **Tidyverse** by Hadley Wickham ([R for Data Science](https://r4ds.hadley.nz")).
Answers on Internet, such as Stack Overflow, might use either of them.
The answer which can solve your problem is the best one for you.

```{r}
#install.packages("tidyverse")
library(tidyverse)

```

**install.packages()** is used to download packages archived in [CRAN](https://cran.r-project.org/). But you do not need to download packages every time you use R, just [once for all]{.ul}.  
**library()** helps load package into your working environment. You need to load your packages [every time you start your RStudio]{.ul}.  
**Tidyverse** package includes a lot of packages, but you still need to download some packages due to your special needs.

### Set Working Directory

```{r}
setwd("/Users/bb")
getwd()
```
[Space and characters other than alphabet]{.ul} might cause problems.  
This step is not necessary, but can save you some time when you need to write your path or file name for multiple times.

### Read Your Data

Tidyverse already includes **readr** package.  
**haven** package can help read other data in formats of "SPSS", "Stata" and "SAS".  
**readxl** package can help read *.xlsx* and *.xls*. 

```{r}
#read_csv() # for ,-seperated csv file
#read_csv2() # for ;-seperated csv file, since you may have , in numbers
#read_dta()
#read_xlxs()
#read_xls()
#etc., etc.
```

You need to save the data you read into global environment. It has nothing to do with the original data file saved your disk.
```{r}
#data <- data_csv()
#data_or_any_name_you_like <- data_csv()
#bb <- data_csv()
#harry_potter <- data_csv()
#darth_vader <- data_csv()
```
The name refers to the table in the environment. If you give it to another table, then the previous one just disappears.

```{r}
data <- starwars
data
data <- penguins
data
```

"<-" is the same as "=", which means "give the value from the right side to the left side".
```{r}
data <- starwars
data
data = penguins
data
```

Encoding is usually causing problem when reading files containing foreign languages. You can fix this problem by **locale()** option in **read_csv()** to set your encoding.
```{r}
#read_csv(locale = "_Your Encoding_")
```

### Read Data from Multiple Files
```{r}
list.files()
list.files(recursive = TRUE) # to read all files in this folder, including those in the folders under this folder.
```
You may want to select the wanted files by their format, such as .csv. Then you need **str_detect()** to find them.
```{r}
str_detect("data_some_survey.csv", "\\.csv")
str_detect("rstudio.exe", "\\..csv")
c("data_some_survey.csv","rstudio.exe","r.exe","long_name_file.somethingelse")
list = c("data_some_survey.csv","rstudio.exe","r.exe","long_name_file.somethingelse")
str_detect(list,"\\.csv")
list[str_detect(list,"\\.csv")]
```


**map()** is basically a way of loop. It runs a certain function to every element in the given list.
```{r}
#list = list.files()
#map(list, read_csv)
```
And you may try to combine them into one table once you have cleaned them.


## Go for Tidy Data
### Tidy Data

THREE LAWS OF TIDY DATA:  
1. Every row is one observation, your unit of analysis.  
2. Every column is one variable.  
3. One table is one dataset.

It is not a golden rule for every software.

```{r}
starwars
```

Every time of observation has its own row for [panel data]{.ul}.

```{r}
#crime4
as_tibble(crime4)
```

### Data Type
```{r echo=FALSE}
data = starwars %>% 
  mutate(gender = as.factor(gender))
glimpse(data)
```
- character*<cha>* for words.
- integer*<int>* for whole numbers.
- double*<dbl>* for numbers with decimal points (fraction).
- factor*<fct>* for categories. [You must change your category data into this type, or they are just characters or numeric values, which are the worst case]{.ul}.

### Pipes
%>% or |> is an operator who passes the output of the current step as an object to the next step.
```{r}
data = starwars %>% 
  mutate(gender = as.factor(gender))
# mutate(
#   .data,
#   ...,
#   .by = NULL,
#   .keep = c("all", "used", "unused", "none"),
#   .before = NULL,
#   .after = NULL
# )
glimpse(data)
```

### Logical Operators &,|

```{r}
TRUE & FALSE
TRUE & TRUE
FALSE & FALSE
TRUE | FALSE
TRUE | TRUE
FALSE | FALSE
```

### Rational Operators >,<,==,!=

```{r}
2 > 1
2 < 1
1 == 2
1 == 1
1 != 0
1 != 1
```


### mutate(), select(), summary()
- **mutate()** can generate new variables or change the value of a certain variable.
- **select()** can select the variables you need (for the next step), those not selected would not be included in the next step.
- **summary()** can give some basic distribution information for the object you input (a dataset or several variables). [It is a good habit to use summary to check missing values or wrong values]{.ul}.

```{r}
data = starwars %>% 
  mutate(sex = as.factor(sex))
summary(data$sex)
data %>% 
  mutate(sex = as.factor(if_else(sex == "male", 1, 0))) %>% 
  select(sex) %>% 
  summary()
data %>% 
  mutate(sex = case_when(sex == "male" ~ 1,
                         sex == "female" ~ 0,
                         .default = -1)) %>% 
  mutate(sex = as.factor(sex)) %>% 
  select(sex) %>% 
  summary()
```

### Change value type by as.factor(), as.numeric()
```{r}
starwars %>% 
  mutate(hair_color = as.factor(hair_color)) %>% 
  select(hair_color) %>% 
  summary()

unique(starwars$hair_color)
```




